### 2023-02-27 products/manuscript/Manuscript.qmd

---
title: "Manuscript for MADA Project"
subtitle: ""
author: Jacob Matta
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

# Abstract

# Introduction

# Methods

# Results

# References

# Discussion