### 2023-03-07 code/processing_code/processingcode2.qmd

---
title: "Processing script 2"
author: "Jacob Matta"
date: "2023-01-03"
output: html_document
---
# Processing script 2: Understanding the rawdata and data source

## Determining the count of residents in GA with at least one dose of the covid-19 vaccines

The count of residents in GA with at least one dose of the Covid-19 vaccines as of Feb 26th, 2023 is 6,935,491. This was found from the Georgia Department of [Public Health Vaccine Distribution Dashboard](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c) on Feb 26th/

## This processing script will look to see how the raw data used for this analysis compares to the overall number of 6,935,491 as of Feb 26th, 2023.

The overall sum of residents in GA with at least one dose of the covid-19 vaccines as of Feb 26th, 2023 should equal 6,935,491

### Load needed packages. make sure they are installed.

```{r}
library(readxl) #for loading Excel files
library(here) #used for setting a path 
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

### Raw Data loading

```{r}
rawdata_location = here("data","raw_data","Feb_26.xlsx")
Feb_26 <- read_excel(rawdata_location)
```

### Codebook loading

```{r}
#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Cleaning

```{r}
#dropping columns that i don't need
#keeping county name to group data together 
Feb_26 <- Feb_26[-c(1:3)]

#removing all of the population data 
Feb_26 = Feb_26[-c(3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33, 35, 37, 39, 41, 43, 45, 47, 49, 51, 53, 55, 57, 59, 61, 63, 65, 67, 69, 71, 73, 75, 77, 79, 81, 83, 85, 87, 89, 91, 93, 95, 97, 99, 101, 103, 105, 107, 109, 111, 113, 115, 117, 119, 121, 123, 125, 127, 129, 131, 133, 135, 137, 139, 141, 143, 145, 147, 149, 151, 153, 155, 157, 159, 161, 163, 165, 167, 169, 171, 173, 175, 177, 179, 181.)]
Feb_26[-c(1)]

#removing the negative values 
Feb_26 [Feb_26 < 0] <- 0

```

### Summing the vaccinations in each column

```{r}
Feb_26_sum_by_col = colSums(Feb_26[ , c(2:91)])     
Feb_26_sum_by_col
```

### Turning the sums by column into a tibble that puts all of the sums of each column into one column

```{r}
Feb_26_sum_tib = tibble(Feb_26_sum_by_col)
Feb_26_sum_tib
```

### Summing all of the vaccinations together

```{r}
Overall_sum = sum(Feb_26_sum_tib) 
Overall_sum
```

### Thoughts

The overall total of residents in GA with at least one dose of the covid-19 vaccines is 5,654,635 which is more than 1 million should of the total about of GA residents with at least one dose of a Covid-19 vaccine that the Georgia Department of Public Health Vaccine Distribution Dashboard reported on Feb, 26th, 2023.

The data that I am using for this project possibly excludes certain vaccines brands (eg. Pfizer, Moderna, Johnson & Johnson) whereas the overall vaccination count represented on the website includes all vaccination brands. This is a question I will attempt to figure out. For now, it is my understanding that there are 5,654,635 Ga residents by census tract who currently have one dose of the Covid-19 vaccine. The brands of Covid-19 vaccines included in the census tract data has yet to be specified.

### Explanation for difference Covid-19 Vaccination totals

1.  The data I downloaded is by census tract

2.  The data displayed on the Georgia Department of [Public Health Vaccine Distribution Dashboard](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c) on Feb 26th is by county level which does not align with census tracts

3.  Going forward for details regarding the raw data you should go to arcGIS insights dashboard for Covid-19 vaccinations among GA Residents with at least one dose found [here](https://insights.arcgis.com/#/view/08d2c1f995524053b1f9669dd0b0b995)

4.  The data dashboard for Covid-19 vaccinations among GA Residents with at least one dose (link above) states that the data downloaded by census tract will not add up to the county level and state level data

### Moving forward
If the vaccinations are not rates (eg. per 100,000) why are there negative values and why does it state that the where "-1 is indicated, either the numerator is less than five or the denominator is less than 25".
-   What is the numerator and what is the denominator?