### 2023-03-19 products/supplement.qmd

# Overview

This file describes all files in this project, the reproducibility procedure for this project, any addition information regarding the methods, results and discussion for this project.

# Description of Project Files

-   `gitignore` file: Tells Git which files and directories to ignore when you make a commit

-   `JM-MADA-Project.Rproj` file: Establishes file paths for project

-   `data` folder: Contains the `processed_data` and `raw_data` folders

    -   `processed_data` folder: Contains the data that has been processed by the processing code

        -   `Sept` folder: Contains processed data for September

            -   `processed_data_31701_Sept.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Sept.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Sept.rds`: Contains processed data for the 31707 zip code

        -   `Oct` folder: Contains processed data for October

            -   `processed_data_31701_Oct.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Oct.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Oct.rds`: Contains processed data for the 31707 zip code

        -   `Nov` folder: Contains processed data for November

            -   `processed_data_31701_Nov.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Nov.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Nov.rds`: Contains processed data for the 31707 zip code

        -   `Dec` folder: Contains processed data for December

            -   `processed_data_31701_Dec.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Dec.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Dec.rds`: Contains processed data for the 31707 zip code

        -   `Jan` folder: Contains processed data for January

            -   `processed_data_31701_Jan.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Jan.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Jan.rds`: Contains processed data for the 31707 zip code

        -   `Feb` folder: Contains processed data for February

            -   `processed_data_31701_Feb.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Feb.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Feb.rds`: Contains processed data for the 31707 zip code

    -   `raw_data` folder: Contains the raw data for this project

        -   `Codebook.xlsx` : Contains the codebook for the raw data

        -   `Sept_17.xlsx`: Contains raw data that was downloaded on September 17th, 2022

        -   `Oct_28.xlsx`: Contains raw data that was downloaded on October 28th, 2022

        -   `Nov_24.xlsx`: Contains raw data that was downloaded on November 24th, 2022

        -   `Dec_29.xlsx`: Contains raw data that was downloaded on December 29th, 2022

        -   `Jan_26.xlsx`: Contains raw data that was downloaded on January 26th, 2023

        -   `Feb_26.xlsx`: Contains raw data that was downloaded on February 26th, 2023

-   `code` folder: Contains `analysis_code` folder and `processing_code` folder which are used to process the raw data into processed data through a series of cleaning steps and then using the analysis code to analyze the data

    -   `processing_code`: Contains the code for data processing in both .R files and .qmd files

        -   `processingcode1.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in September 2022

        -   `processingcode1.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in September 2022

        -   `processingcode2.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in October 2022

        -   `processingcode2.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in October 2022

        -   `processingcode3.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in November 2022

        -   `processingcode3.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in October 2022

        -   `processingcode4.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in December 2022

        -   `processingcode4.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in December 2022

        -   `processingcode5.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in January 2023

        -   `processingcode5.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in January 2022

        -   `processingcode6.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in February 2023

        -   `processingcode6.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in February 2023

        -   `readme.md`: provides a brief description of the files in `processing_code` folder

    -   `analysis_code`: Contains the code for the analysis

        -   `analysiscode1.R`: Contains the code an analysis of the total amount of African Americans with at least one COVID-19 vaccine in the 31701 zip code
        -   `readme.md`: provides a brief description of the analysis files

-   `products` folder: Contains files for the manuscript and supplementary material of this project

    -   `apa.csl` file: APA format for citations

    -   `Manuscript.qmd` file: Contains the manuscript for this project
    -   `My_Library.bib`: Bibliography for this project

    -   `readme.md`: Contains a description of the files in the `products` folder

    -   `supplement.qmd`: Contains supplementart material relevent to reproducing this project and a brief description of every file in this project

-   `results` folder: