### 2023-04-05 code/analysis_code/analysiscode2.qmd

---
title: "Total Amount of Vaccinations by Gender Among Black population in the Target Zip Codes"
output: 
  html_document
echo: FALSE 
---

# Library

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr)
library(ggplot2)
library(reshape2) #reshaping data from wide to long format
library(plotly) #creating an interactive line plot 
```

# Total Amount of vaccinations by gender among Black population in the target zip codes

------------------------------------------------------------------------

## Zip code: 31701

### Loading the `processed_combined_31701.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31701", "processed_combined_31701.rds")
processed_combined_31701 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31701

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31701_Gender_black = processed_combined_31701 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31701_total_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31701_total_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31701_total_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31701_total_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31701_total_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31701_total_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31701_total_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31701_total_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31701_total_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31701_total_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31701_total_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31701_total_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(3371, 3358, 3368, 3379, 3387, 3397),
                          Female_Total = c(4426, 4422, 4387, 4394, 4405, 4420))
```

### Converting the Gender_Sum_by_month_31701 to long data in order to plot
```{r}
Gender_Sum_by_month_31701_long = reshape2::melt(Gender_Sum_by_month_31701, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31701_long <- Gender_Sum_by_month_31701_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31701 zip code

```{r}
plot_Gender_black_vax = Gender_Sum_by_month_31701_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Total_black_vax = plot_Total_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31701") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax = ggplotly(plot_Gender_black_vax)
plotly_plot_Gender_black_vax
```

------------------------------------------------------------------------

## Zip code: 31705

### Loading the `processed_combined_31705.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31705", "processed_combined_31705.rds")
processed_combined_31705 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31705

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31705_total_black = processed_combined_31705 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31705_total_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31705_total_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31705_total_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31705_total_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31705_total_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31705_total_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31705_total_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31705_total_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31705_total_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31705_total_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31705_total_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31705_total_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(4394, 4397, 4399, 4391, 4403, 4399),
                          Female_Total = c(5989, 5930, 5904, 5910, 5934, 5949))
```

### Converting the Gender_Sum_by_month_31705 to long data in order to plot
```{r}
Gender_Sum_by_month_31705_long = reshape2::melt(Gender_Sum_by_month_31705, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31705_long <- Gender_Sum_by_month_31705_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31705 zip code

```{r}
plot_Gender_black_vax_31705 = Gender_Sum_by_month_31705_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Total_black_vax = plot_Total_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31705") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax_31705
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax_31705 = ggplotly(plot_Gender_black_vax_31705)
plotly_plot_Gender_black_vax_31705
```

------------------------------------------------------------------------

## Zip code: 31707

### Loading the `processed_combined_31707.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31707", "processed_combined_31707.rds")
processed_combined_31707 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31707

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31707_total_black = processed_combined_31707 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31707_total_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31707_total_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31707_total_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31707_total_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31707_total_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31707_total_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31707_total_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31707_total_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31707_total_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31707_total_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31707_total_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31707_total_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(3698, 3674, 3678, 3680, 3684, 3681),
                          Female_Total = c(5337, 5311, 5319, 5325, 5351, 5361))
```

### Converting the Gender_Sum_by_month_31707 to long data in order to plot
```{r}
Gender_Sum_by_month_31707_long = reshape2::melt(Gender_Sum_by_month_31707, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31707_long <- Gender_Sum_by_month_31707_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31707 zip code

```{r}
plot_Gender_black_vax_31707 = Gender_Sum_by_month_31707_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Total_black_vax = plot_Total_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31707") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax_31707
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax_31707 = ggplotly(plot_Gender_black_vax_31707)
plotly_plot_Gender_black_vax_31707
```

# Thoughts regarding exploratory analysis 
In each zip code there is a large gap between females that are Black and males that are Black. I am pretty taken back by how large the gap is.