### 2023-04-07 code/analysis_code/exploratoryanalysis.qmd
Part 4

---
title: "Total Amount of Vaccinations Among Black population in the Target Zip Codes"
output: 
  html_document
echo: FALSE 
---

# Library

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr)
library(ggplot2)
library(reshape2) #reshaping data from wide to long format
library(plotly) #creating an interactive line plot 
```

# Analysis of the Amount of Vaccinations Among Black population in Each Target Zip Code

## Zip code: 31701

### Loading the processed data in order to begin analysis for the 31701 zip code

```{r}
# Loading September 31701
data_location1 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Sept.rds")
processed_data_31701_Sept = readRDS(data_location1)

# Loading October 31701 and joining september and october 
data_location2 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Oct.rds")
processed_data_31701_Oct = readRDS(data_location2)
processed_combined <- left_join(processed_data_31701_Sept, processed_data_31701_Oct, by = c("FIPS" = "FIPS"))

# Loading November 31701
data_location3 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Nov.rds")
processed_data_31701_Nov = readRDS(data_location3)
processed_combined <- left_join(processed_combined, processed_data_31701_Nov, by = c("FIPS" = "FIPS"))

# Loading December 31701
data_location4 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Dec.rds")
processed_data_31701_Dec = readRDS(data_location4)
processed_combined <- left_join(processed_combined, processed_data_31701_Dec, by = c("FIPS" = "FIPS"))

# Loading January 31701
data_location5 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Jan.rds")
processed_data_31701_Jan = readRDS(data_location5)
processed_combined <- left_join(processed_combined, processed_data_31701_Jan, by = c("FIPS" = "FIPS"))

# Loading February 31701
data_location6 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Feb.rds")
processed_data_31701_Feb = readRDS(data_location6)
processed_combined_31701 <- left_join(processed_combined, processed_data_31701_Feb, by = c("FIPS" = "FIPS"))
```

### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender

```{r}
save_data_location = here("data", "processed_data","Zip_31701", "processed_combined_31701.rds")
saveRDS(processed_combined_31701, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31701 zip code

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31701_total_black = processed_combined_31701 %>% select(starts_with("total_black_vax"))
```

### Summing each Month variable by census tract

```{r}
#Summing each column for each month 
sum(processed_combined_31701_total_black$total_black_vax_Sept_17) 
sum(processed_combined_31701_total_black$total_black_vax_Oct_28) 
sum(processed_combined_31701_total_black$total_black_vax_Nov_24) 
sum(processed_combined_31701_total_black$total_black_vax_Dec_29) 
sum(processed_combined_31701_total_black$total_black_vax_Jan_26) 
sum(processed_combined_31701_total_black$total_black_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Total = c(7797, 7780, 7755, 7773, 7792, 7817)) 
```

### gglot2: Creating a line plot for the vaccination totals by month for the African American population in the 31701 zip code

```{r}
plot_31701_total_black_vax = Sum_by_month_31701 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot 
plot_31701_total_black_vax = plot_31701_total_black_vax + 
  ggtitle("Vaccination Totals for Black population in the 31701 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
plot_31701_total_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_31701_total_black_vax = ggplotly(plot_31701_total_black_vax)
plotly_31701_total_black_vax
```

## Zip code: 31705

### Loading the processed data in order to begin analysis for the 31705 zip code

```{r}
#Loading September 31705
data_location7 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Sept.rds")
processed_data_31705_Sept = readRDS(data_location7)

# Loading October 31705 and joining september and october 
data_location8 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Oct.rds")
processed_data_31705_Oct = readRDS(data_location8)
processed_combined <- left_join(processed_data_31705_Sept, processed_data_31705_Oct, by = c("FIPS" = "FIPS"))

# Loading November 31705
data_location9 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Nov.rds")
processed_data_31705_Nov = readRDS(data_location9)
processed_combined <- left_join(processed_combined, processed_data_31705_Nov, by = c("FIPS" = "FIPS"))

# Loading December 31705
data_location10= here("data", "processed_data", "Zip_31705", "processed_data_31705_Dec.rds")
processed_data_31705_Dec = readRDS(data_location10)
processed_combined <- left_join(processed_combined, processed_data_31705_Dec, by = c("FIPS" = "FIPS"))

# Loading January 31705
data_location11 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Jan.rds")
processed_data_31705_Jan = readRDS(data_location11)
processed_combined <- left_join(processed_combined, processed_data_31705_Jan, by = c("FIPS" = "FIPS"))

# Loading February 31705
data_location12 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Feb.rds")
processed_data_31705_Feb = readRDS(data_location12)
processed_combined_31705 <- left_join(processed_combined, processed_data_31705_Feb, by = c("FIPS" = "FIPS"))

```

### Saving the combined data for Black population for future use

```{r}
save_data_location = here("data", "processed_data","Zip_31705", "processed_combined_31705.rds")
saveRDS(processed_combined_31705, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31705 zip code

```{r}
processed_combined_31705_total_black = processed_combined_31705 %>% select(starts_with("total_black_vax"))
```

### Summing each column for each month

```{r}

sum(processed_combined_31705_total_black$total_black_vax_Sept_17)

sum(processed_combined_31705_total_black$total_black_vax_Oct_28)

sum(processed_combined_31705_total_black$total_black_vax_Nov_24)

sum(processed_combined_31705_total_black$total_black_vax_Dec_29)

sum(processed_combined_31705_total_black$total_black_vax_Jan_26)

sum(processed_combined_31705_total_black$total_black_vax_Feb_26)
```

### Separating sums by month into a data fram consisting of two variables: Date and Total

### Date contains months September-February

### Total contains the vaccination totals for each month

```{r}
Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                                Total = c(10383, 10327, 10303, 10301, 10337, 10348)) 
```

### ggplot creating a line plot for the vaccination totals by month for the 31705 zip code African American population

```{r}
plot_31705_total_black_vax = Sum_by_month_31705 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot 
plot_31705_total_black_vax = plot_31705_total_black_vax + 
  ggtitle("Vaccination Totals for Black population in the 31705 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
```

## Zip code: 31707

### Loading the processed data in order to begin analysis for the 31707 zip code

```{r}
# Loading September 31707
data_location13 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Sept.rds")
processed_data_31707_Sept = readRDS(data_location13)

# Loading October 31707 and joining september and october 
data_location14 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Oct.rds")
processed_data_31707_Oct = readRDS(data_location14)
processed_combined <- left_join(processed_data_31707_Sept, processed_data_31707_Oct, by = c("FIPS" = "FIPS"))

# Loading November 31707
data_location15 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Nov.rds")
processed_data_31707_Nov = readRDS(data_location15)
processed_combined <- left_join(processed_combined, processed_data_31707_Nov, by = c("FIPS" = "FIPS"))

# Loading December 31707
data_location16= here("data", "processed_data", "Zip_31707", "processed_data_31707_Dec.rds")
processed_data_31707_Dec = readRDS(data_location16)
processed_combined <- left_join(processed_combined, processed_data_31707_Dec, by = c("FIPS" = "FIPS"))

# Loading January 31707
data_location17 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Jan.rds")
processed_data_31707_Jan = readRDS(data_location17)
processed_combined <- left_join(processed_combined, processed_data_31707_Jan, by = c("FIPS" = "FIPS"))

# Loading February 31707
data_location18 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Feb.rds")
processed_data_31707_Feb = readRDS(data_location18)
processed_combined_31707 <- left_join(processed_combined, processed_data_31707_Feb, by = c("FIPS" = "FIPS"))
```

### Saving the combined data for Black population for future use

```{r}
save_data_location = here("data", "processed_data","Zip_31707", "processed_combined_31707.rds")
saveRDS(processed_combined_31707, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31707 zip code

```{r}
processed_combined_31707_total_black = processed_combined_31707 %>% select(starts_with("total_black_vax"))

```

### Summing each column for each month

```{r}
sum(processed_combined_31707_total_black$total_black_vax_Sept_17) 
sum(processed_combined_31707_total_black$total_black_vax_Oct_28) 
sum(processed_combined_31707_total_black$total_black_vax_Nov_24) 
sum(processed_combined_31707_total_black$total_black_vax_Dec_29) 
sum(processed_combined_31707_total_black$total_black_vax_Jan_26) 
sum(processed_combined_31707_total_black$total_black_vax_Feb_26)
```

### Separating sums by month into a data fram consisting of two variables: Date and Total

### Date contains months September-February

### Total contains the vaccination totals for each month

```{r}
Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                                Total = c(9035, 8985, 8997, 9005, 9035, 9042))

```

###ggplot: creating a line plot for the vaccination totals by month for the 31707 zip code African American population

```{r}
plot_31707_total_black_vax = Sum_by_month_31707 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot
plot_31707_total_black_vax = plot_31707_total_black_vax + 
  ggtitle("Vaccination Totals for Black population in the 31707 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
```

## Comparing plots the 3 target zip codes

### Taking a look at each of the line plots for the different zip codes and comparing the total vaccinations by month in the African American population

```{r}
plot_31701_total_black_vax
plot_31705_total_black_vax
plot_31707_total_black_vax
```

### Combining all plots into one

```{r}
## Creating a new dataframe called "Total_black_vax" by combining the vaccination totals from 31701 and 3105 by using leftjoin() by the Date variable 
Total_black_vax = left_join(Sum_by_month_31701, Sum_by_month_31705, by = c("Date" = "Date"))

```

### Combining the new dataframe "Total_black_vax" with the 31707 zip vaccination totals so all three target zip codes are in the same data frame

```{r}
Total_black_vax = left_join(Total_black_vax, Sum_by_month_31707, by = c("Date" = "Date"))
```

### Renaming column totals to match the zip

### Haing to do this because the the total variable names are Total.x, Total.y and Total

### Need to specificy each total variable by zip using the following code

```{r}
Total_black_vax <- Total_black_vax %>% 
  rename("31701" = "Total.x") %>% 
  rename("31705" = "Total.y") %>%
  rename("31707" = "Total")
```

### Converting data to long format in order to create a line plot using ggplot

```{r}
Total_black_vax_long = reshape2::melt(Total_black_vax, id = "Date")
```

### Changing the variable name "variable" to "Zip_code" and "value" to "Vaccination_total" for clarity

### Doing this because when the data was converted to a long format the column names we changed to "variable" which refers to the zip code and "value" which refers to the amount of vaccinations

### Changing "variable" to "Zip_code" and "value" to "Vaccination_total"

```{r}
Total_black_vax_long <- Total_black_vax_long %>% 
  rename("Zip_code" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### ggplot: Making a line graph that compares the vaccination totals for each African American population in each zip code

```{r}
plot_Total_black_vax = Total_black_vax_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Zip_code, 
                                color = Zip_code)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Total_black_vax = plot_Total_black_vax + 
  ggtitle("Vaccination Totals for Black population in Target Zip Codes") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Total_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_Total_black_vax = ggplotly(plot_Total_black_vax)
plotly_Total_black_vax
```

# Total Amount of vaccinations by Gender Among Black Population in the Target Zip Codes

## Zip code: 31701

### Loading the `processed_combined_31701.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31701", "processed_combined_31701.rds")
processed_combined_31701 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31701

#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31701_Gender_black = processed_combined_31701 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(3371, 3358, 3368, 3379, 3387, 3397),
                          Female_Total = c(4426, 4422, 4387, 4394, 4405, 4420))
```

### Converting the Gender_Sum_by_month_31701 to long data in order to plot
```{r}
Gender_Sum_by_month_31701_long = reshape2::melt(Gender_Sum_by_month_31701, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31701_long <- Gender_Sum_by_month_31701_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31701 zip code

```{r}
plot_Gender_black_vax = Gender_Sum_by_month_31701_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Gender_black_vax = plot_Gender_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31701") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax = ggplotly(plot_Gender_black_vax)
plotly_plot_Gender_black_vax
```

## Zip code: 31705

### Loading the `processed_combined_31705.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31705", "processed_combined_31705.rds")
processed_combined_31705 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31705

#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31705_Gender_black = processed_combined_31705 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(4394, 4397, 4399, 4391, 4403, 4399),
                          Female_Total = c(5989, 5930, 5904, 5910, 5934, 5949))
```

### Converting the Gender_Sum_by_month_31705 to long data in order to plot
```{r}
Gender_Sum_by_month_31705_long = reshape2::melt(Gender_Sum_by_month_31705, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31705_long <- Gender_Sum_by_month_31705_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31705 zip code

```{r}
plot_Gender_black_vax_31705 = Gender_Sum_by_month_31705_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Gender_black_vax = plot_Gender_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31705") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax_31705
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax_31705 = ggplotly(plot_Gender_black_vax_31705)
plotly_plot_Gender_black_vax_31705
```

## Zip code: 31707

### Loading the `processed_combined_31707.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31707", "processed_combined_31707.rds")
processed_combined_31707 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31707

#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31707_Gender_black = processed_combined_31707 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(3698, 3674, 3678, 3680, 3684, 3681),
                          Female_Total = c(5337, 5311, 5319, 5325, 5351, 5361))
```

### Converting the Gender_Sum_by_month_31707 to long data in order to plot
```{r}
Gender_Sum_by_month_31707_long = reshape2::melt(Gender_Sum_by_month_31707, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31707_long <- Gender_Sum_by_month_31707_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31707 zip code

```{r}
plot_Gender_black_vax_31707 = Gender_Sum_by_month_31707_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Gender_black_vax = plot_Gender_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31707") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax_31707
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax_31707 = ggplotly(plot_Gender_black_vax_31707)
plotly_plot_Gender_black_vax_31707
```

# Thoughts regarding exploratory analysis 
In each zip code there is a large gap between females that are Black and males that are Black. I am pretty taken back by how large the gap is.

# Total Amount of vaccinations by Age among Black population in the target zip codes

## Zip code: 31701

### Loading the `processed_combined_31701.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31701", "processed_combined_31701.rds")
processed_combined_31701 = readRDS(save_data_location)
```

### Dropping all variables except for black age groups within 31701

#### Using the select() function to subset the data frame by columns that start with an age group

```{r}
processed_combined_31701_Age_black = processed_combined_31701 %>% 
  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
```

### Summing each Month variable by age group by month
```{r}
#Summing for ages 5-9 by month 
sum(processed_combined_31701_Age_black$total_05_09_vax_Sept_17) 
sum(processed_combined_31701_Age_black$total_05_09_vax_Oct_28) 
sum(processed_combined_31701_Age_black$total_05_09_vax_Nov_24) 
sum(processed_combined_31701_Age_black$total_05_09_vax_Dec_29) 
sum(processed_combined_31701_Age_black$total_05_09_vax_Jan_26) 
sum(processed_combined_31701_Age_black$total_05_09_vax_Feb_26)

#Summing for ages 10-17 by month 
sum(processed_combined_31701_Age_black$total_10_17_vax_Sept_17) 
sum(processed_combined_31701_Age_black$total_10_17_vax_Oct_28) 
sum(processed_combined_31701_Age_black$total_10_17_vax_Nov_24) 
sum(processed_combined_31701_Age_black$total_10_17_vax_Dec_29) 
sum(processed_combined_31701_Age_black$total_10_17_vax_Jan_26) 
sum(processed_combined_31701_Age_black$total_10_17_vax_Feb_26)

#Summing for ages 18-44 by month 
sum(processed_combined_31701_Age_black$total_18_44_vax_Sept_17) 
sum(processed_combined_31701_Age_black$total_18_44_vax_Oct_28) 
sum(processed_combined_31701_Age_black$total_18_44_vax_Nov_24) 
sum(processed_combined_31701_Age_black$total_18_44_vax_Dec_29) 
sum(processed_combined_31701_Age_black$total_18_44_vax_Jan_26) 
sum(processed_combined_31701_Age_black$total_18_44_vax_Feb_26)

#Summing for ages 45-64 by month 
sum(processed_combined_31701_Age_black$total_45_64_vax_Sept_17) 
sum(processed_combined_31701_Age_black$total_45_64_vax_Oct_28) 
sum(processed_combined_31701_Age_black$total_45_64_vax_Nov_24) 
sum(processed_combined_31701_Age_black$total_45_64_vax_Dec_29) 
sum(processed_combined_31701_Age_black$total_45_64_vax_Jan_26) 
sum(processed_combined_31701_Age_black$total_45_64_vax_Feb_26)

#Summing for ages 65+ by month 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Sept_17) 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Oct_28) 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Nov_24) 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Dec_29) 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Jan_26) 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Feb_26)

```

### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Age_Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Age_05_09 = c(123, 126, 126, 127, 127, 128),
                          Age_10_17 = c(642, 635, 633, 637, 630, 629),
                          Age_18_44 = c(3310, 3316, 3313, 3323, 3325, 3331),
                          Age_45_64 = c(3229, 3244, 3239, 3259, 3264, 3258),
                          Age_65Plus = c(2576, 2570, 2569, 2574, 2568, 2572))

```

### Converting the Age_Sum_by_month_31701 to long data in order to plot
```{r}
Age_Sum_by_month_31701_long = reshape2::melt(Age_Sum_by_month_31701, id = "Date")
```

### Changing "variable" to "Age" and "value" to "Vaccination_total"

```{r}
Age_Sum_by_month_31701_long <- Age_Sum_by_month_31701_long %>% 
  rename("Age" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 31701 zip code

```{r}
plot_Age_black_vax = Age_Sum_by_month_31701_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Age, 
                                color = Age)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Age_black_vax = plot_Age_black_vax + 
  ggtitle("Vaccination Totals for Black population by Age in 31701") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Age_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
plotly_plot_Age_black_vax
```

------------------------------------------------------------------------
## Zip code: 31705

### Loading the `processed_combined_31705.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31705", "processed_combined_31705.rds")
processed_combined_31705 = readRDS(save_data_location)
```

### Dropping all variables except for black age groups within 31705

#### Using the select() function to subset the data frame by columns that start with an age group

```{r}
processed_combined_31705_Age_black = processed_combined_31705 %>% 
  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
```

### Summing each Month variable by age group by month
```{r}
#Summing for ages 5-9 by month 
sum(processed_combined_31705_Age_black$total_05_09_vax_Sept_17) 
sum(processed_combined_31705_Age_black$total_05_09_vax_Oct_28) 
sum(processed_combined_31705_Age_black$total_05_09_vax_Nov_24) 
sum(processed_combined_31705_Age_black$total_05_09_vax_Dec_29) 
sum(processed_combined_31705_Age_black$total_05_09_vax_Jan_26) 
sum(processed_combined_31705_Age_black$total_05_09_vax_Feb_26)

#Summing for ages 10-17 by month 
sum(processed_combined_31705_Age_black$total_10_17_vax_Sept_17) 
sum(processed_combined_31705_Age_black$total_10_17_vax_Oct_28) 
sum(processed_combined_31705_Age_black$total_10_17_vax_Nov_24) 
sum(processed_combined_31705_Age_black$total_10_17_vax_Dec_29) 
sum(processed_combined_31705_Age_black$total_10_17_vax_Jan_26) 
sum(processed_combined_31705_Age_black$total_10_17_vax_Feb_26)

#Summing for ages 18-44 by month 
sum(processed_combined_31705_Age_black$total_18_44_vax_Sept_17) 
sum(processed_combined_31705_Age_black$total_18_44_vax_Oct_28) 
sum(processed_combined_31705_Age_black$total_18_44_vax_Nov_24) 
sum(processed_combined_31705_Age_black$total_18_44_vax_Dec_29) 
sum(processed_combined_31705_Age_black$total_18_44_vax_Jan_26) 
sum(processed_combined_31705_Age_black$total_18_44_vax_Feb_26)

#Summing for ages 45-64 by month 
sum(processed_combined_31705_Age_black$total_45_64_vax_Sept_17) 
sum(processed_combined_31705_Age_black$total_45_64_vax_Oct_28) 
sum(processed_combined_31705_Age_black$total_45_64_vax_Nov_24) 
sum(processed_combined_31705_Age_black$total_45_64_vax_Dec_29) 
sum(processed_combined_31705_Age_black$total_45_64_vax_Jan_26) 
sum(processed_combined_31705_Age_black$total_45_64_vax_Feb_26)

#Summing for ages 65+ by month 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Sept_17) 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Oct_28) 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Nov_24) 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Dec_29) 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Jan_26) 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Feb_26)

```

### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February 
### Total contains the vaccination totals for each month

```{r}
Age_Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Age_05_09 = c(210, 211, 209, 201, 202, 202),
                          Age_10_17 = c(1159, 1171, 1169, 1163, 1168, 1167),
                          Age_18_44 = c(5201, 5215, 5239, 5263, 5278, 5281), #pretty notable increase for ages 18-44 
                          Age_45_64 = c(4380, 4381, 4366, 4364, 4367, 4367),
                          Age_65Plus = c(2983, 3000, 2997, 2997, 2994, 2989))

```

### Converting the Age_Sum_by_month_31705 to long data in order to plot
```{r}
Age_Sum_by_month_31705_long = reshape2::melt(Age_Sum_by_month_31705, id = "Date")
```

### Changing "variable" to "Age" and "value" to "Vaccination_total"

```{r}
Age_Sum_by_month_31705_long <- Age_Sum_by_month_31705_long %>% 
  rename("Age" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 31705 zip code

```{r}
plot_Age_black_vax = Age_Sum_by_month_31705_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Age, 
                                color = Age)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Age_black_vax = plot_Age_black_vax + 
  ggtitle("Vaccination Totals for Black population by Age in 31705") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Age_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
plotly_plot_Age_black_vax
```

------------------------------------------------------------------------

## Zip code: 31707

### Loading the `processed_combined_31707.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31707", "processed_combined_31707.rds")
processed_combined_31707 = readRDS(save_data_location)
```

### Dropping all variables except for black age groups within 31707

#### Using the select() function to subset the data frame by columns that start with an age group

```{r}
processed_combined_31707_Age_black = processed_combined_31707 %>% 
  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
```

### Summing each Month variable by age group by month
```{r}
#Summing for ages 5-9 by month 
sum(processed_combined_31707_Age_black$total_05_09_vax_Sept_17) 
sum(processed_combined_31707_Age_black$total_05_09_vax_Oct_28) 
sum(processed_combined_31707_Age_black$total_05_09_vax_Nov_24) 
sum(processed_combined_31707_Age_black$total_05_09_vax_Dec_29) 
sum(processed_combined_31707_Age_black$total_05_09_vax_Jan_26) 
sum(processed_combined_31707_Age_black$total_05_09_vax_Feb_26)

#Summing for ages 10-17 by month 
sum(processed_combined_31707_Age_black$total_10_17_vax_Sept_17) 
sum(processed_combined_31707_Age_black$total_10_17_vax_Oct_28) 
sum(processed_combined_31707_Age_black$total_10_17_vax_Nov_24) 
sum(processed_combined_31707_Age_black$total_10_17_vax_Dec_29) 
sum(processed_combined_31707_Age_black$total_10_17_vax_Jan_26) 
sum(processed_combined_31707_Age_black$total_10_17_vax_Feb_26)

#Summing for ages 18-44 by month 
sum(processed_combined_31707_Age_black$total_18_44_vax_Sept_17) 
sum(processed_combined_31707_Age_black$total_18_44_vax_Oct_28) 
sum(processed_combined_31707_Age_black$total_18_44_vax_Nov_24) 
sum(processed_combined_31707_Age_black$total_18_44_vax_Dec_29) 
sum(processed_combined_31707_Age_black$total_18_44_vax_Jan_26) 
sum(processed_combined_31707_Age_black$total_18_44_vax_Feb_26)

#Summing for ages 45-64 by month 
sum(processed_combined_31707_Age_black$total_45_64_vax_Sept_17) 
sum(processed_combined_31707_Age_black$total_45_64_vax_Oct_28) 
sum(processed_combined_31707_Age_black$total_45_64_vax_Nov_24) 
sum(processed_combined_31707_Age_black$total_45_64_vax_Dec_29) 
sum(processed_combined_31707_Age_black$total_45_64_vax_Jan_26) 
sum(processed_combined_31707_Age_black$total_45_64_vax_Feb_26)

#Summing for ages 65+ by month 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Sept_17) 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Oct_28) 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Nov_24) 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Dec_29) 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Jan_26) 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Feb_26)

```

# The Amount of Vaccinations Among Black population in the Control Zip Code

## Zip code: 30815

### Loading the processed data in order to begin analysis for the 30815 zip code

```{r}
# Loading September 30815
data_location1 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Sept.rds")
processed_data_30815_Sept = readRDS(data_location1)

# Loading October 30815 and joining september and october 
data_location2 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Oct.rds")
processed_data_30815_Oct = readRDS(data_location2)
processed_combined <- left_join(processed_data_30815_Sept, processed_data_30815_Oct, by = c("FIPS" = "FIPS"))

# Loading November 30815
data_location3 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Nov.rds")
processed_data_30815_Nov = readRDS(data_location3)
processed_combined <- left_join(processed_combined, processed_data_30815_Nov, by = c("FIPS" = "FIPS"))

# Loading December 30815
data_location4 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Dec.rds")
processed_data_30815_Dec = readRDS(data_location4)
processed_combined <- left_join(processed_combined, processed_data_30815_Dec, by = c("FIPS" = "FIPS"))

# Loading January 30815
data_location5 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Jan.rds")
processed_data_30815_Jan = readRDS(data_location5)
processed_combined <- left_join(processed_combined, processed_data_30815_Jan, by = c("FIPS" = "FIPS"))

# Loading February 30815
data_location6 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Feb.rds")
processed_data_30815_Feb = readRDS(data_location6)
processed_combined_30815 <- left_join(processed_combined, processed_data_30815_Feb, by = c("FIPS" = "FIPS"))
```

### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender

```{r}
save_data_location = here("data", "processed_data","Zip_30815", "processed_combined_30815.rds")
saveRDS(processed_combined_30815, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 30815 zip code

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_30815_total_black = processed_combined_30815 %>% select(starts_with("total_black_vax"))
```

### Summing each Month variable by census tract

```{r}
#Summing each column for each month 
sum(processed_combined_30815_total_black$total_black_vax_Sept_17) 
sum(processed_combined_30815_total_black$total_black_vax_Oct_28) 
sum(processed_combined_30815_total_black$total_black_vax_Nov_24) 
sum(processed_combined_30815_total_black$total_black_vax_Dec_29) 
sum(processed_combined_30815_total_black$total_black_vax_Jan_26) 
sum(processed_combined_30815_total_black$total_black_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Total = c(8758, 8688, 8639, 8593, 8623, 8666)) 
```

### gglot2: Creating a line plot for the vaccination totals by month for the African American population in the 30815 zip code

```{r}
plot_30815_total_black_vax = Sum_by_month_30815 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot 
plot_30815_total_black_vax = plot_30815_total_black_vax + 
  ggtitle("Vaccination Totals for Black population in the 30815 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
plot_30815_total_black_vax
```
### Turning ggplot into an interactive plot

```{r}
plotly_30815_total_black_vax = ggplotly(plot_30815_total_black_vax)
plotly_30815_total_black_vax
```

# The Amount of Vaccinations by Gender Among Black Population in the Control Zip Code

## Zip code: 30815

### Loading the `processed_combined_30815.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_30815", "processed_combined_30815.rds")
processed_combined_30815 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 30815

#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_30815_Gender_black = processed_combined_30815 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Sept_17) 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Oct_28) 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Nov_24) 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Dec_29) 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Jan_26) 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Sept_17) 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Oct_28) 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Nov_24) 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Dec_29) 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Jan_26) 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(3526, 3514, 3503, 3484, 3489, 3498),
                          Female_Total = c(5232, 5174, 5136, 5109, 5134, 5168))
```

### Converting the Gender_Sum_by_month_30815 to long data in order to plot
```{r}
Gender_Sum_by_month_30815_long = reshape2::melt(Gender_Sum_by_month_30815, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_30815_long <- Gender_Sum_by_month_30815_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 30815 zip code

```{r}
plot_Gender_black_vax = Gender_Sum_by_month_30815_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Gender_black_vax = plot_Gender_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 30815") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax = ggplotly(plot_Gender_black_vax)
plotly_plot_Gender_black_vax
```

# Total Amount of vaccinations by Age among Black population in the Control Zip Code

## Zip code: 30815

### Loading the `processed_combined_30815.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_30815", "processed_combined_30815.rds")
processed_combined_30815 = readRDS(save_data_location)
```

### Dropping all variables except for black age groups within 30815

#### Using the select() function to subset the data frame by columns that start with an age group

```{r}
processed_combined_30815_Age_black = processed_combined_30815 %>% 
  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
```

### Summing each Month variable by age group by month
```{r}
#Summing for ages 5-9 by month 
sum(processed_combined_30815_Age_black$total_05_09_vax_Sept_17) 
sum(processed_combined_30815_Age_black$total_05_09_vax_Oct_28) 
sum(processed_combined_30815_Age_black$total_05_09_vax_Nov_24) 
sum(processed_combined_30815_Age_black$total_05_09_vax_Dec_29) 
sum(processed_combined_30815_Age_black$total_05_09_vax_Jan_26) 
sum(processed_combined_30815_Age_black$total_05_09_vax_Feb_26)

#Summing for ages 10-17 by month 
sum(processed_combined_30815_Age_black$total_10_17_vax_Sept_17) 
sum(processed_combined_30815_Age_black$total_10_17_vax_Oct_28) 
sum(processed_combined_30815_Age_black$total_10_17_vax_Nov_24) 
sum(processed_combined_30815_Age_black$total_10_17_vax_Dec_29) 
sum(processed_combined_30815_Age_black$total_10_17_vax_Jan_26) 
sum(processed_combined_30815_Age_black$total_10_17_vax_Feb_26)

#Summing for ages 18-44 by month 
sum(processed_combined_30815_Age_black$total_18_44_vax_Sept_17) 
sum(processed_combined_30815_Age_black$total_18_44_vax_Oct_28) 
sum(processed_combined_30815_Age_black$total_18_44_vax_Nov_24) 
sum(processed_combined_30815_Age_black$total_18_44_vax_Dec_29) 
sum(processed_combined_30815_Age_black$total_18_44_vax_Jan_26) 
sum(processed_combined_30815_Age_black$total_18_44_vax_Feb_26)

#Summing for ages 45-64 by month 
sum(processed_combined_30815_Age_black$total_45_64_vax_Sept_17) 
sum(processed_combined_30815_Age_black$total_45_64_vax_Oct_28) 
sum(processed_combined_30815_Age_black$total_45_64_vax_Nov_24) 
sum(processed_combined_30815_Age_black$total_45_64_vax_Dec_29) 
sum(processed_combined_30815_Age_black$total_45_64_vax_Jan_26) 
sum(processed_combined_30815_Age_black$total_45_64_vax_Feb_26)

#Summing for ages 65+ by month 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Sept_17) 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Oct_28) 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Nov_24) 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Dec_29) 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Jan_26) 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Feb_26)

```

### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Age_Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Age_05_09 = c(201, 213, 211, 215, 211, 211),
                          Age_10_17 = c(1116, 1113, 1115, 1118, 1115, 1114),
                          Age_18_44 = c(5001, 5009, 5020, 5026, 5025, 5032),
                          Age_45_64 = c(5026, 5017, 5020, 5009, 5013, 5020),
                          Age_65Plus = c(2816, 2825, 2827, 2814, 2828, 2832))

```

### Converting the Age_Sum_by_month_30815 to long data in order to plot
```{r}
Age_Sum_by_month_30815_long = reshape2::melt(Age_Sum_by_month_30815, id = "Date")
```

### Changing "variable" to "Age" and "value" to "Vaccination_total"

```{r}
Age_Sum_by_month_30815_long <- Age_Sum_by_month_30815_long %>% 
  rename("Age" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 30815 zip code

```{r}
plot_Age_black_vax = Age_Sum_by_month_30815_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Age, 
                                color = Age)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Age_black_vax = plot_Age_black_vax + 
  ggtitle("Vaccination Totals for Black population by Age in 30815") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Age_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
plotly_plot_Age_black_vax
```