### 2023-04-18 code/processing_code/UntitledQMD2.qmd
Population sums by zip

# Processing Section for Control Zip Code

## Processing 1: Getting Population Totals by Control Zip Code 30815 in the Month of September

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Sept_17.xlsx")
Sept_17 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Sept <- filter(Sept_17, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Sept [Richmond_POP_Sept < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_POP_Sept = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Sept <- Richmond_POP_Sept  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Sept<- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Richmond_POP_Sept$total_asian_POP <- Richmond_POP_Sept$total_masian_POP + Richmond_POP_Sept$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Sept$total_black_POP <- Richmond_POP_Sept$total_mblack_POP + Richmond_POP_Sept$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Sept$total_orace_POP <- Richmond_POP_Sept$total_morace_POP + Richmond_POP_Sept$total_forace_POP

#POP total for White Residents
Richmond_POP_Sept$total_white_POP <- Richmond_POP_Sept$total_mwhite_POP + Richmond_POP_Sept$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Sept_sums <- Richmond_POP_Sept[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
#All of the following FIPS or census tracts correspond to the 30815 zip code
Zip_30815_POP_Sept <-filter(Richmond_POP_Sept_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
```
Not that the new data set `Zip_30815_POP_Sept` only has 7 observations because the orinigal data set only had 7 of the 13 census tracts found within the zip code 30815

### Change the column names for 30815 Sept
```{r}
colnames(Zip_30815_POP_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17") 
```

### Deleting excess environment objects and saving data as processed_data_POP_sept
```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Sept,Sept_17, codebook, Richmond_POP_Sept_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30815", "processed_data_POP_30815_Sept.rds")

saveRDS(Zip_30815_POP_Sept, save_data_location4)
```

## Processing 2: Getting Population Totals by Target Zips 30815 for the Month of October

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Oct_28.xlsx")
Oct_28 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Oct <- filter(Oct_28, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Oct [Richmond_POP_Oct < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_POP_Oct = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Oct <- Richmond_POP_Oct  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Oct<- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.

```{r}
#POP total Asian Residents
Richmond_POP_Oct$total_asian_POP <- Richmond_POP_Oct$total_masian_POP + Richmond_POP_Oct$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Oct$total_black_POP <- Richmond_POP_Oct$total_mblack_POP + Richmond_POP_Oct$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Oct$total_orace_POP <- Richmond_POP_Oct$total_morace_POP + Richmond_POP_Oct$total_forace_POP

#POP total for White Residents
Richmond_POP_Oct$total_white_POP <- Richmond_POP_Oct$total_mwhite_POP + Richmond_POP_Oct$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Oct_sums <- Richmond_POP_Oct[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
#All of the following FIPS or census tracts correspond to the 30815 zip code

Zip_30815_POP_Oct <-filter(Richmond_POP_Oct_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
```

### Change the column names for 30815 Oct

```{r}
colnames(Zip_30815_POP_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28") 
```

### Deleting excess environment objects and saving data as processed_data_POP_Oct

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Oct,Oct_28, codebook, Richmond_POP_Oct_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Oct.rds")

saveRDS(Zip_30815_POP_Oct, file = save_data_location4)
```

## Processing 3: : Getting Population Totals by Target Zips 30815 for the Month of November

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Nov_24.xlsx")
Nov_24 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Nov_24)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Nov <- filter(Nov_24, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Nov [Richmond_POP_Nov < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_POP_Nov = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Nov <- Richmond_POP_Nov  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Nov<- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.

```{r}
#POP total Asian Residents
Richmond_POP_Nov$total_asian_POP <- Richmond_POP_Nov$total_masian_POP + Richmond_POP_Nov$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Nov$total_black_POP <- Richmond_POP_Nov$total_mblack_POP + Richmond_POP_Nov$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Nov$total_orace_POP <- Richmond_POP_Nov$total_morace_POP + Richmond_POP_Nov$total_forace_POP

#POP total for White Residents
Richmond_POP_Nov$total_white_POP <- Richmond_POP_Nov$total_mwhite_POP + Richmond_POP_Nov$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Nov_sums <- Richmond_POP_Nov[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
#All of the following FIPS or census tracts correspond to the 30815 zip code

Zip_30815_POP_Nov <-filter(Richmond_POP_Nov_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")

```

### Change the column names for 30815 Nov

```{r}
colnames(Zip_30815_POP_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Nov_24", "total_male_POP_Nov_24", "total_female_POP_Nov_24", "total_masian_POP_Nov_24", "total_mblack_POP_Nov_24", "total_morace_POP_Nov_24", "total_mwhite_POP_Nov_24", "total_fasian_POP_Nov_24", "total_fblack_POP_Nov_24", "total_forace_POP_Nov_24", "total_fwhite_POP_Nov_24", "total_asian_POP_Nov_24","total_black_POP_Nov_24", "total_orace_POP_Nov_24", "total_white_POP_Nov_24", "total_05_09_POP_Nov_24", "total_10_17_POP_Nov_24", "total_18_44_POP_Nov_24", "total_45_64_POP_Nov_24", "total_65Plus_POP_Nov_24") 
```

### Deleting excess environment objects and saving data as processed_data_POP_Nov

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Nov,Nov_24, codebook, Richmond_POP_Nov_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Nov.rds")

saveRDS(Zip_30815_POP_Nov, file = save_data_location4)

```

## Processing 4: Getting Population Totals by Target Zips 30815 for the Month of December

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Dec_29.xlsx")
Dec_29 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Dec_29)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Dec <- filter(Dec_29, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Dec [Richmond_POP_Dec < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_POP_Dec = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Dec <- Richmond_POP_Dec  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Dec<- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.

```{r}
#POP total Asian Residents
Richmond_POP_Dec$total_asian_POP <- Richmond_POP_Dec$total_masian_POP + Richmond_POP_Dec$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Dec$total_black_POP <- Richmond_POP_Dec$total_mblack_POP + Richmond_POP_Dec$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Dec$total_orace_POP <- Richmond_POP_Dec$total_morace_POP + Richmond_POP_Dec$total_forace_POP

#POP total for White Residents
Richmond_POP_Dec$total_white_POP <- Richmond_POP_Dec$total_mwhite_POP + Richmond_POP_Dec$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Dec_sums <- Richmond_POP_Dec[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
Zip_30815_POP_Dec <-filter(Richmond_POP_Dec_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")

```

### Change the column names for 30815 zip code data

```{r}
colnames(Zip_30815_POP_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Dec_29", "total_male_POP_Dec_29", "total_female_POP_Dec_29", "total_masian_POP_Dec_29", "total_mblack_POP_Dec_29", "total_morace_POP_Dec_29", "total_mwhite_POP_Dec_29", "total_fasian_POP_Dec_29", "total_fblack_POP_Dec_29", "total_forace_POP_Dec_29", "total_fwhite_POP_Dec_29", "total_asian_POP_Dec_29","total_black_POP_Dec_29", "total_orace_POP_Dec_29", "total_white_POP_Dec_29", "total_05_09_POP_Dec_29", "total_10_17_POP_Dec_29", "total_18_44_POP_Dec_29", "total_45_64_POP_Dec_29", "total_65Plus_POP_Dec_29")

```

### Deleting excess environment objects and saving data as processed_data_POP_Dec

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Dec,Dec_29, codebook, Richmond_POP_Dec_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Dec.rds")

saveRDS(Zip_30815_POP_Dec, file = save_data_location4)
```

## Processing 5: Getting Population Totals by Target Zips 30815 for the Month of January

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Jan_26.xlsx")
Jan_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Jan_26)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Jan <- filter(Jan_26, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Jan [Richmond_POP_Jan < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_POP_Jan = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Jan <- Richmond_POP_Jan  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Jan<- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.

```{r}
#POP total Asian Residents
Richmond_POP_Jan$total_asian_POP <- Richmond_POP_Jan$total_masian_POP + Richmond_POP_Jan$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Jan$total_black_POP <- Richmond_POP_Jan$total_mblack_POP + Richmond_POP_Jan$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Jan$total_orace_POP <- Richmond_POP_Jan$total_morace_POP + Richmond_POP_Jan$total_forace_POP

#POP total for White Residents
Richmond_POP_Jan$total_white_POP <- Richmond_POP_Jan$total_mwhite_POP + Richmond_POP_Jan$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Jan_sums <- Richmond_POP_Jan[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
Zip_30815_POP_Jan <-filter(Richmond_POP_Jan_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")

```

### Change the column names for 30815 Jan

```{r}
colnames(Zip_30815_POP_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Jan_26", "total_male_POP_Jan_26", "total_female_POP_Jan_26", "total_masian_POP_Jan_26", "total_mblack_POP_Jan_26", "total_morace_POP_Jan_26", "total_mwhite_POP_Jan_26", "total_fasian_POP_Jan_26", "total_fblack_POP_Jan_26", "total_forace_POP_Jan_26", "total_fwhite_POP_Jan_26", "total_asian_POP_Jan_26","total_black_POP_Jan_26", "total_orace_POP_Jan_26", "total_white_POP_Jan_26", "total_05_09_POP_Jan_26", "total_10_17_POP_Jan_26", "total_18_44_POP_Jan_26", "total_45_64_POP_Jan_26", "total_65Plus_POP_Jan_26") 
```

### Deleting excess environment objects and saving data as processed_data_POP_Jan

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Jan,Jan_26, codebook, Richmond_POP_Jan_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Jan.rds")

saveRDS(Zip_30815_POP_Jan, file = save_data_location4)
```

## Processing 6: : Getting Population Totals by Target Zips 30815 for the Month of February

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Feb_26.xlsx")
Feb_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Feb_26)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Feb <- filter(Feb_26, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Feb [Richmond_POP_Feb < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_POP_Feb = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Feb <- Richmond_POP_Feb  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Feb<- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.

```{r}
#POP total Asian Residents
Richmond_POP_Feb$total_asian_POP <- Richmond_POP_Feb$total_masian_POP + Richmond_POP_Feb$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Feb$total_black_POP <- Richmond_POP_Feb$total_mblack_POP + Richmond_POP_Feb$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Feb$total_orace_POP <- Richmond_POP_Feb$total_morace_POP + Richmond_POP_Feb$total_forace_POP

#POP total for White Residents
Richmond_POP_Feb$total_white_POP <- Richmond_POP_Feb$total_mwhite_POP + Richmond_POP_Feb$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Feb_sums <- Richmond_POP_Feb[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
Zip_30815_POP_Feb <-filter(Richmond_POP_Feb_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")

```

### Change the column names for 30815 Feb

```{r}
colnames(Zip_30815_POP_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Feb_26", "total_male_POP_Feb_26", "total_female_POP_Feb_26", "total_masian_POP_Feb_26", "total_mblack_POP_Feb_26", "total_morace_POP_Feb_26", "total_mwhite_POP_Feb_26", "total_fasian_POP_Feb_26", "total_fblack_POP_Feb_26", "total_forace_POP_Feb_26", "total_fwhite_POP_Feb_26", "total_asian_POP_Feb_26","total_black_POP_Feb_26", "total_orace_POP_Feb_26", "total_white_POP_Feb_26", "total_05_09_POP_Feb_26", "total_10_17_POP_Feb_26", "total_18_44_POP_Feb_26", "total_45_64_POP_Feb_26", "total_65Plus_POP_Feb_26") 
```

### Deleting excess environment objects and saving data as processed_data_POP_Feb

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Feb,Feb_26, codebook, Richmond_POP_Feb_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Feb.rds")

saveRDS(Zip_30815_POP_Feb, file = save_data_location4)
```

# Processing Section for Control Zip Code: 30907

## Processing 1: Getting Population Totals by Control Zip Code 30907 in the Month of September

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Sept_17.xlsx")
Sept_17 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Sept <-filter(Sept_17, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Sept)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Sept [Zip_30907_POP_Sept < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_POP_Sept = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Sept <- Zip_30907_POP_Sept  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Sept<- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Sept$total_asian_POP <- Zip_30907_POP_Sept$total_masian_POP + Zip_30907_POP_Sept$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Sept$total_black_POP <- Zip_30907_POP_Sept$total_mblack_POP + Zip_30907_POP_Sept$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Sept$total_orace_POP <- Zip_30907_POP_Sept$total_morace_POP + Zip_30907_POP_Sept$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Sept$total_white_POP <- Zip_30907_POP_Sept$total_mwhite_POP + Zip_30907_POP_Sept$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Sept_sums <- Zip_30907_POP_Sept[,c(1:4,185:204)]  
```

### Change the column names for 30907 Sept
```{r}
colnames(Zip_30907_POP_Sept_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Sept.rds")

saveRDS(Zip_30907_POP_Sept_sums, save_data_location4)
```

## Processing 2: Getting Population Totals by Control Zip Code 30907 in the Month of October

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Oct_28.xlsx")
Oct_28 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Oct <-filter(Oct_28, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Oct)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Oct [Zip_30907_POP_Oct < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_POP_Oct = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Oct <- Zip_30907_POP_Oct  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Oct<- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Oct$total_asian_POP <- Zip_30907_POP_Oct$total_masian_POP + Zip_30907_POP_Oct$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Oct$total_black_POP <- Zip_30907_POP_Oct$total_mblack_POP + Zip_30907_POP_Oct$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Oct$total_orace_POP <- Zip_30907_POP_Oct$total_morace_POP + Zip_30907_POP_Oct$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Oct$total_white_POP <- Zip_30907_POP_Oct$total_mwhite_POP + Zip_30907_POP_Oct$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Oct_sums <- Zip_30907_POP_Oct[,c(1:4,185:204)]  
```

### Change the column names for 30907 Oct
```{r}
colnames(Zip_30907_POP_Oct_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Oct.rds")

saveRDS(Zip_30907_POP_Oct_sums, save_data_location4)
```

## Processing 3: : Getting Population Totals by Control Zip Code 30907 in the Month of November

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Nov_24.xlsx")
Nov_24 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Nov <-filter(Nov_24, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Nov)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Nov [Zip_30907_POP_Nov < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_POP_Nov = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Nov <- Zip_30907_POP_Nov  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Nov<- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Nov$total_asian_POP <- Zip_30907_POP_Nov$total_masian_POP + Zip_30907_POP_Nov$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Nov$total_black_POP <- Zip_30907_POP_Nov$total_mblack_POP + Zip_30907_POP_Nov$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Nov$total_orace_POP <- Zip_30907_POP_Nov$total_morace_POP + Zip_30907_POP_Nov$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Nov$total_white_POP <- Zip_30907_POP_Nov$total_mwhite_POP + Zip_30907_POP_Nov$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Nov_sums <- Zip_30907_POP_Nov[,c(1:4,185:204)]  
```

### Change the column names for 30907 Nov
```{r}
colnames(Zip_30907_POP_Nov_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Nov_24", "total_male_POP_Nov_24", "total_female_POP_Nov_24", "total_masian_POP_Nov_24", "total_mblack_POP_Nov_24", "total_morace_POP_Nov_24", "total_mwhite_POP_Nov_24", "total_fasian_POP_Nov_24", "total_fblack_POP_Nov_24", "total_forace_POP_Nov_24", "total_fwhite_POP_Nov_24", "total_asian_POP_Nov_24","total_black_POP_Nov_24", "total_orace_POP_Nov_24", "total_white_POP_Nov_24", "total_05_09_POP_Nov_24", "total_10_17_POP_Nov_24", "total_18_44_POP_Nov_24", "total_45_64_POP_Nov_24", "total_65Plus_POP_Nov_24") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Nov.rds")

saveRDS(Zip_30907_POP_Nov_sums, save_data_location4)
```

## Processing 4: : Getting Population Totals by Control Zip Code 30907 in the Month of December

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Dec_29.xlsx")
Dec_29 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Dec <-filter(Dec_29, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Dec)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Dec [Zip_30907_POP_Dec < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_POP_Dec = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Dec <- Zip_30907_POP_Dec  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Dec<- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Dec$total_asian_POP <- Zip_30907_POP_Dec$total_masian_POP + Zip_30907_POP_Dec$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Dec$total_black_POP <- Zip_30907_POP_Dec$total_mblack_POP + Zip_30907_POP_Dec$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Dec$total_orace_POP <- Zip_30907_POP_Dec$total_morace_POP + Zip_30907_POP_Dec$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Dec$total_white_POP <- Zip_30907_POP_Dec$total_mwhite_POP + Zip_30907_POP_Dec$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Dec_sums <- Zip_30907_POP_Dec[,c(1:4,185:204)]  
```

### Change the column names for 30907 Dec
```{r}
colnames(Zip_30907_POP_Dec_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Dec_29", "total_male_POP_Dec_29", "total_female_POP_Dec_29", "total_masian_POP_Dec_29", "total_mblack_POP_Dec_29", "total_morace_POP_Dec_29", "total_mwhite_POP_Dec_29", "total_fasian_POP_Dec_29", "total_fblack_POP_Dec_29", "total_forace_POP_Dec_29", "total_fwhite_POP_Dec_29", "total_asian_POP_Dec_29","total_black_POP_Dec_29", "total_orace_POP_Dec_29", "total_white_POP_Dec_29", "total_05_09_POP_Dec_29", "total_10_17_POP_Dec_29", "total_18_44_POP_Dec_29", "total_45_64_POP_Dec_29", "total_65Plus_POP_Dec_29") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Dec.rds")

saveRDS(Zip_30907_POP_Dec_sums, save_data_location4)
```

## Processing 5: : Getting Population Totals by Control Zip Code 30907 in the Month of January

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Jan_26.xlsx")
Jan_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Jan <-filter(Jan_26, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Jan)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Jan [Zip_30907_POP_Jan < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_POP_Jan = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Jan <- Zip_30907_POP_Jan  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Jan<- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Jan$total_asian_POP <- Zip_30907_POP_Jan$total_masian_POP + Zip_30907_POP_Jan$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Jan$total_black_POP <- Zip_30907_POP_Jan$total_mblack_POP + Zip_30907_POP_Jan$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Jan$total_orace_POP <- Zip_30907_POP_Jan$total_morace_POP + Zip_30907_POP_Jan$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Jan$total_white_POP <- Zip_30907_POP_Jan$total_mwhite_POP + Zip_30907_POP_Jan$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Jan_sums <- Zip_30907_POP_Jan[,c(1:4,185:204)]  
```

### Change the column names for 30907 Jan
```{r}
colnames(Zip_30907_POP_Jan_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Jan_26", "total_male_POP_Jan_26", "total_female_POP_Jan_26", "total_masian_POP_Jan_26", "total_mblack_POP_Jan_26", "total_morace_POP_Jan_26", "total_mwhite_POP_Jan_26", "total_fasian_POP_Jan_26", "total_fblack_POP_Jan_26", "total_forace_POP_Jan_26", "total_fwhite_POP_Jan_26", "total_asian_POP_Jan_26","total_black_POP_Jan_26", "total_orace_POP_Jan_26", "total_white_POP_Jan_26", "total_05_09_POP_Jan_26", "total_10_17_POP_Jan_26", "total_18_44_POP_Jan_26", "total_45_64_POP_Jan_26", "total_65Plus_POP_Jan_26") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Jan.rds")

saveRDS(Zip_30907_POP_Jan_sums, save_data_location4)
```

## Processing 6: : Getting Population Totals by Control Zip Code 30907 in the Month of February

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Feb_26.xlsx")
Feb_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Feb <-filter(Feb_26, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Feb)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Feb [Zip_30907_POP_Feb < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_POP_Feb = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Feb <- Zip_30907_POP_Feb  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Feb<- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Feb$total_asian_POP <- Zip_30907_POP_Feb$total_masian_POP + Zip_30907_POP_Feb$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Feb$total_black_POP <- Zip_30907_POP_Feb$total_mblack_POP + Zip_30907_POP_Feb$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Feb$total_orace_POP <- Zip_30907_POP_Feb$total_morace_POP + Zip_30907_POP_Feb$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Feb$total_white_POP <- Zip_30907_POP_Feb$total_mwhite_POP + Zip_30907_POP_Feb$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Feb_sums <- Zip_30907_POP_Feb[,c(1:4,185:204)]  
```

### Change the column names for 30907 Feb
```{r}
colnames(Zip_30907_POP_Feb_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Feb_26", "total_male_POP_Feb_26", "total_female_POP_Feb_26", "total_masian_POP_Feb_26", "total_mblack_POP_Feb_26", "total_morace_POP_Feb_26", "total_mwhite_POP_Feb_26", "total_fasian_POP_Feb_26", "total_fblack_POP_Feb_26", "total_forace_POP_Feb_26", "total_fwhite_POP_Feb_26", "total_asian_POP_Feb_26","total_black_POP_Feb_26", "total_orace_POP_Feb_26", "total_white_POP_Feb_26", "total_05_09_POP_Feb_26", "total_10_17_POP_Feb_26", "total_18_44_POP_Feb_26", "total_45_64_POP_Feb_26", "total_65Plus_POP_Feb_26") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Feb.rds")

saveRDS(Zip_30907_POP_Feb_sums, save_data_location4)
```