### 2023-04-27 Hatfield Review.qmd
Kelly Review

---
title: Project Review  
author: Kelly Hatfield
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project:  COVID-19 Vaccination in Albany, GA

Name of project author(s): Jacob Matta

Name of project reviewer: Kelly Hatfield

### Feedback and Comments
 The background and introduction are  thorough and proide good motiviation to the importance of the study.

The abstract could use help with clarity, perhaps reframe the study some: 
The outcome of interest is the change in the proportion of residents vacccinated over time from September - February. The predictor of interest is the COVID-19 vaccine hesitancy program intervention. It would be helpful also to help contextualize the three zip codes of interest.

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The hypotheis and questions can be more explicitly stated in the manuscript. For example, in the objectives section you mention data cleaning methods. The objective should be stated more succintly and explicitly:

We aim to assess any trend in vaccine uptake among our targeted community during the time of health literacy programs (September - February 2022)

Your hypothesis is a little unclear to me too. For example, you are looking for an increase in the number of Albany citizens who receive the vaccine in the time period. However, this is a value that can't decrease. So are you looking for an increase in the trend (interrupted tiem series), and increase in the daily rate among unvaccinated (where your denominator will change daily), an increase in the zip codes of interest versus other zip codes where targeted interventions werent applied? The objective/hypothesis might need to be fleshed out a bit more.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data structure is well explained and sources described. It is meddled in the with the importation and cleaning. I would rather encourage the author to separate the findings a bit.

### Summary assessment
* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Data cleaning process is well described. Similar to comment above, it is interwoven with description of the data sources. The manuscript would benefit from separate paragraphs discussing these elements and some editorial review (e.g., ensuring run-on sentences are shortened, fragments removed). Some mentions of "data were cleaned" without explaining exaclty how. An example would be:

"For this analysis, we decided to exclude purina body condition scores as it was implemented later in the study, and most dogs do not have that information. The remaining columns were cleaned to make the data more consistent and in the format required for statistical analysis. "

could be streamlined to: 
"We excluded condition scores because they were not consistently collected and had high numbers of missing values. We recoded XXXX variables from XXX to XXX to facilitate easier statistical analysis.""

### Summary assessment

* some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

### Summary assessment
* defensible but not optimal analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

### Summary assessment
* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments
Discussion not present at time of review.

### Summary assessment

* strong, complete and clear discussion

## Further comments

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

### Summary assessment

* strong level of thorougness

## Further comments
---